### 2012-04-12 ScalingPerformance.tex

\section{Scaling Performances}

Real time versus size, bottle neck, extended pp, trials on scinet and ranger, 
p3dfft,