### 2012-04-17 Initial.tex
Significant changes to Poisson Solver, precised the skeleton, added a few discussion shere and there...

\subsection{Initial Conditions}
\label{subsec:init}

As mention in section \ref{sec:structure}, the code start off by reading a set of initial conditions.
These correspond to a  $6 \times N$ phase-space array, where $N$ is the number of particles in the
local node. Although most applications so far were based on initial conditions that follow Gaussian statistics,
we have developed a non-Gaussian initial condition generator that we briefly describe in this section. 
{\bf Vincent, here you go!}