### 2012-04-17 PID.tex

\subsection{Particle identification tags}
\label{PID}

A system of particle identification can be turned on, which basically allows to track each particle's trajectory
between checkpoints. Such a tool is useful to a number of applications, from reconstruction of halo merging history to {\bf (what else?)}
The particle tag system has been implemented as an array of double integers, {\tt PID}, 
and assigns a unique integer to each particle during the initialization stage. The location of the tag on the {\tt PID} array 
matches the location of the corresponding particle on the {\tt xv} array, hence it acts as if the latter array had an extra dimension.
The location change only when particles exist the local volume, in which case the tag is sent along with the particle in the {\tt pass\_particle} subroutine.
Deleted particles results in deleted flags, and the {\tt PID} array gets written to file at each particle checkpoint.
{\bf (Anything else to say here?)}