### 2012-04-17 Halos.tex
More improvements

\section{Runtime Halo Finder}
\label{sec:halo}

Spherical overdensities, search algorithm, 
provided halo information, halo bias, comparison to PS and ST.
{\bf (Ilian, You can lead the way here...)}