### 2012-05-01 Conclusion.tex
Upgraded extendedpp.tex, plus minor corrections everywhere.

\section{Conclusion}

The code is publicly available on github.com under {\tt cubep3m}, and extra documentation about the structure, 
compiling and running strategy is can be found at {\tt www.wiki.cita.utoronto.ca/mediawiki/index.php/CubePM}.

\section*{Acknowledgements}