### 2012-05-01 ScalingPerformance.tex

\section{Scaling Performances}

Real time versus size, bottle neck, description of the largest runs, etc.
{\bf (Ilian, I will need your input here as well. Hugh, do you have anything you would like to share 
concerning the largest runs you did on Blue Gene? )}