### 2012-06-11 Halos.tex
included predictions for bias

\section{Runtime Halo Finder}
\label{sec:halo}

%Spherical overdensities, search algorithm, 
%provided halo information, halo bias, comparison to PS and ST.
%{\bf (Ilian, You can lead the way here...)}

We have implemented a halo finding procedure, which we have developed 
based on the spherical overdensity (SO) approach \citep{1994MNRAS.271..676L}.
In the interest of speed and efficiency the halo catalogues are constructed 
on-the-fly at a pre-determined list of redshifts. The halo finding is 
massively-parallel and threaded based on the main {\small CUBEP3M} data structures 
discussed in section \ref{sec:structure}. The code first builds the 
fine-mesh density for each sub-domain using CIC or NGP interpolation. It then 
proceeds to search for and record all local density maxima above certain
threshold (typically set to 100 above mean density) within the local 
sub-domain's physical volume (excluding the sub-domain buffer zones). It then 
uses parabolic interpolation on the density field to determine more precisely
the location of the maximum within the densest cell, and records the peak 
position and value. The halo centre determined this way agrees closely with 
the centre-of-mass of the halo particles discussed below.

Once the list of peak positions is generated, they are sorted from the highest 
to the lowest density value. Then each of the halo candidates is inspected 
independently, starting with the highest peak. The grid mass is accumulated 
in spherical shells of fine grid cells surrounding the maximum, until the 
mean density within the halo drops below a pre-defined overdensity cutoff 
(usually set to 178 in units of the mean, in accordance to the top-hat 
collapse model). As we accumulate mass we remove it from the mesh, so that no 
mass element is double-counted. This method is thus inappropriate for finding 
sub-halos as within this framework those are naturally incorporated in their 
host halos. Because the halos are found on a grid of finite-size cells, it is 
possible, especially for the low-mass halos, to overshoot the target overdensity.
When this occurs we use an analytical halo density profile to correct the 
halo mass and radius to the values corresponding to the target overdensity. 
This analytical density profile is given by Truncated Isothermal Sphere (TIS) 
profile \citep{1999MNRAS.307..203S,2001MNRAS.325..468I} for overdensities below 
$\sim130$, and $1/r^2$ for lower overdensities. The density profile has a
similar outer slope (the relevant one here) to the Navarro, Frenk and White 
\citep[NFW][]{1997ApJ...490..493N} profile, but extends to lower overdensities
and matches well the virialization shock position given by the Bertschinger 
self-similar collapse solution \citep{1985ApJS...58...39B}.

Once the correct halo mass, radius and position are determined, we find all 
particles which are within the halo radius. Their positions and velocities are
used to calculate the halo centre-of-mass, bulk velocity, internal velocity 
dispersion and the three angular momentum components, all of which are then 
included in the final halo catalogues. We also calculate the total mass of
all particles within the halo radius, also listed in the halo data. This mass
is very close, but  typically slightly lower, than the halo mass calculated 
based on the gridded density field. The particle centre-of-mass corresponds 
very well to the halo centre found based on the gridded mass distribution.

%Compared to Tinker
%44% low for 50 particles
%22% low for 400 particles
%12% low for 1000 particles

\begin{figure}%[ht]
%  \vskip -0.5cm 
  \begin{center}
    \includegraphics[width=3.2in]{graphs/mf_z0_Tinker.eps}
  \end{center}
  \caption{Simulated halo multiplicity function, 
    $\frac{M^2}{\bar{\rho}}\frac{dn}{dM}$ based on a
    RANGER4000 simulation with $3.2\,h^{-1} \mbox{Gpc}$ box and $4000^3$ 
    particles (solid, red in the online version). For reference we also show a widely-used 
    precise fit by \citet{2008ApJ...688..709T} (blue, dashed). 
    \label{mf}}
\end{figure}

A sample halo mass function produced based on our inlined SO halo finder 
at redshift $z=0$ from a RANGER4000 simulation is shown in Fig. \ref{mf}. We compare our result to the 
precise fit presented recently by \citet{2008ApJ...688..709T}. Unlike most
other widely-used fits like the one by \citet{2002MNRAS.329...61S}, which are based on friends-of-friends (FOF)
halo finders, the \citet{2008ApJ...688..709T} fit is based on the
SO search algorithm, whose masses are systematically different 
from the FOF masses \citep[e.g.][]{2007MNRAS.374....2R,2008ApJ...688..709T}, 
making this fit a better base for comparison here. Results show excellent
agreement, within $\sim10$ per cent for all halos with masses corresponding to
1000 particles or more. Lower-mass halos are somewhat under-counted compared
to the \citet{2008ApJ...688..709T} fit, by $\sim20$ per cent for 400 particles and 
by $\sim40$ per cent for 50 particles. This is due to the grid-based nature of our
SO halo finder, which misses some of the low-mass halos. It was shown, however, that using more sophisticated
halo finders (available through post-processing due to their heavier memory
footprint) allows to recover the expected mass function.

A second test of the accuracy of the halo finder algorithm is to extract the halo power spectrum $P_{h}(k)$ and compare the halo bias
with theoretical predictions. The halo bias is defined as $b(k) = \sqrt{P(k)/P_{h}(k)}$ and is shown in Fig. \ref{fig:halo}.
The results are organized in four mass bins, where the lowest mass consists of 20-200 particles, the second bin 
contains 200-2000 particles and so on. Comparisons with the linear predictions of \cite{Sheth_et_al} shows
results consistent findings of \cite{Thinker}, namely that low-mass bias is over-predicted and high-mass bias is under-predicted.

\begin{figure}%[ht]
%  \vskip -0.5cm 
  \begin{center}
    \includegraphics[width=3.2in]{graphs/bias.eps}
  \end{center}
  \caption{Halo bias, measured at $z=0$ from a RANGER4000 simulation with a side length of $3.2 h^{-1}$Gpc.
  Bottom to top curves correspond to haloes of different masses, from light to massive. The four bins are decades
  of halo masses; the first contains haloes made of 20-200 particles, the second of 200-2,000 and so on.
  The dashed lines are linear predictions from \protect \cite{Sheth_et_al}.
    \label{fig:halo}}
\end{figure}

%{\bf Show some examples and comparisions to analytical fits and other halo finders.}